The effects of priming in human computation

### 2014-05-07 docs/draft/pnas1.tex
work on manuscript: rework intro, add refs

Platforms like Amazon Mechanical Turk (AMT) make it possible to submit 
batches of ``microtasks'' to a large pool of human workers, who do these 
tasks for fun and remuneration.  A typical task might be the flagging of
innapropriate images, something that people are still better at than 
computers.

AMT makes human cognition as fluid a
resource as compute cycles: it is available on demand in small units, with low 
transaction cost, from anywhere via the Internet. The result is that one can 
treat AMT like a compute cluster, and seemlessly integrate human and machine
intelligence in large-scale computations.

This development invites a new perspective on human cognition itself.  Authors 
have put forward the term HPU (Human Processing Unit) likening an AMT worker
to her machine equivalent, the CPU.  Thinking of people as networked 
processors is not new.  Allan Newell argued 
that a prime function human organizations, such as companies or 
governmental institutions, is actually information processing.

If we adopt this view, how do people differ from CPUs as processors?  
Common sense suggests a major difference is that people do not respond 
in a consistent way to given inputs.  Not only do people vary widely from 
one another, but they differ in time due to mood, level of alertness, and 
any number of other incidental factors. One expects this to be especially 
pronounced for high-level, ill-defined tasks---precisely those for which 
people are usually better-suited than computers.

Here we investigate an aspect of the variability of HPUs.  Specifically,
we look at how a person's output during a task is influenced by the completion
of earlier tasks.  In engineered systems, the phenomenon whereby a system's 
output depends not only on current inputs, but also on the history of inputs,
is called \textit{hysteresis}.  Thus, we investigate the hysteretic aspect
of HPUs.
\td{in psychology this is called priming}

We test hysteresis by presenting AMT workers with
five images that depict food with strong cultural overtones, and ask them to
label each image with descriptive words.  Before the workers label the test
images, we expose them to different priming treatments.  We then observe 
whether and how this effects workers' labels, by looking at how
often workers make explicit cultural references, and how specific their labels 
are.

We induce priming in two ways.  The first way consists of presenting 
workers with different `priming images'.  The workers must also label these 
images, and from the worker's perspective no distinction is
made between the test and priming images.  We call this 
priming modality \textit{in-task priming}, because it attempts to measure the
effects that earlier parts of a task have on later ones.

All workers are exposed to 
one of three priming image sets.  The set that we call \textit{ambiguous},
consists of images of food chosen to avoid obvious connection to a 
specific culture.  One could liken this treatment to a control, hovewer, as
we discuss later, the notion of a control treatment in the context of a priming
experiment is problematic.  The two other priming image sets are 
\textit{ingredients}, consisting of images of separated ingredients, and 
\textit{culture}, consisting of images of overt and specific cultural 
activities which do not feature food.

The natural expectation is that workers
shown the cultural images would provide more culture-oriented labels than
either of the other two image sets.  In the case of the ingredients image set,
beyond encouraging food-oriented labels, one might expect focus be drawn to 
the \textit{constituents} of meals, eliciting labels of higher specificity than
for the ambiguous set.

The second way we induce priming is by presenting the workers with the name 
of a (fictitious) research funder.  Workers are shown either the funder name  
``The National Foundation for Nutritional Awarenes'' or ``The Global
Foundation for Cultural Recognition'', or else they are shown no name.  These 
names were
chosen with the intent of biasing the workers to produce, respectively, more
food- and culture-oriented labels.  We call the second priming method 
\textit{framing}.

Framing is intended to pesent a biasing factor somewhat resembling that which 
the double-blind experimental design tries to avoid.  One might expect that, 
given the overt nature of the funder names, which strongly imply a research
intent, framing should have a strong effect.  We include it as reference point,
for comparison to in-task priming.  This allows us
to ask whether altering the first half of the task can have as strong an 
influence as does revealing an intent-laden funder name.

By taking a computational approach to human cognition, likening priming to
hysteresis effects of physical systems, we gain traction on the notion of 
priming itself. An important aspect of our contribution is to suggest
an algorithmic definition of priming based on so-called 
\textit{classifier algorithms}.  This definition makes the notion of priming
rigorous, and provides a universal approach to detecting priming effects which
can be implemented without knowing what the nature of the effects are.

We present this definition in the next section.  Following, we present
our results together with discussion, and describe our methodology last.

@@

\section*{Acknowledgements}
\section*{References}

>>>>

Microtask crowdsourcing platforms like Amazon Mechanical Turk (AMT) make it 
possible to submit batches of small tasks to a large pool of human workers, 
who do the tasks for fun and remuneration.  A typical task might be the 
flagging of innapropriate images, something that people are still better at 
than computers.

Such platforms human cognition as fluid a resource as compute cycles: it is 
available on demand in small units, with low transaction cost, from anywhere 
via the Internet \footnote{\href{http://mturk.com}{http://mturk.com}}.  The 
result is that one can treat AMT like a compute cluster, and seemlessly 
integrate human and machine intelligence in large-scale computations.

This development invites a new perspective on the role of human cognition 
itself in technical systems.  Researchers have put forward the term HPU 
(Human co-Processing Unit)\footnote{We acknowledge the irony of this term:
the original meaning of `computer' was a person who performs computations,
its reference to machines being a personification brought into usage some time
after 1900}, viewing the introduction of 
microtask represents a fundamentally new computing architecture, in the same
way that the GPUs differ from CPUs \cite{5543192}.  Thinking of networked 
people as computate systems is not new.  Working in operations research, Simon 
and Newell held the primary activity of human organizations to be 
computation\cite{Simon:1958fp}.

In adopting this view, we must contend with a stark difference between 
people and CPUs: even when given the same instructions, people do not produce
the same output.  In other words, one pays for the ``high level instructions 
set'' of HPUs by accepting variability in outputs.

Variability raises a quality control issue. There have been many studies 
into how user interface design, payments, social reputation schemes, and the 
disclosure of the wider purpose of tasks affect worker outputs. We will 
review the prior work in the section following this one.
{\em Here, we report our investigation of worker {\em hysteresis}, that is, 
the tendency of a worker's output during one task to be influenced by other 
recently completed tasks}.

We test hysteresis by presenting AMT workers with
five images that depict meals with strong cultural overtones, and ask them to
label each image with descriptive words.  Before presenting test images,
we expose the workers to different priming treatments.  We observe how priming
treatments effect the workers' subsequent choice of labels, by looking at 
how specific the labels are in their meaning, and how often workers make 
cultural references.

We employ two priming modalities in this study.  
In the first modality, which we call \textit{in-task priming}, workers are 
shown one of three sets of priming images.  The \textit{cultural}
image set, \textbf{Fig. 2}, shows images chosen for their clear depiction of a 
iconic cultural scene.  The \textit{ingredients} image set, \textbf{Fig. 3}, 
shows separated food ingredients, as opposed to the prepared meals shown in 
the test set, and avoids strong cultural signal.  The \textit{ambiguous} 
image set contains images of prepared meals, similar to the test set, but 
traces of cultural references are not prominent.  These images are shown right
before the test images, and they must be labeled in the same way.  From the 
workers' view-point there is no distinction between the priming and test 
images.

The second priming modality, which we call \textit{framing}, consists of
presenting the workers with the name of a (fictitious) research funder before
presenting any images to be labeled.  Workers are shown either the funder 
name  
``The National Foundation for Nutritional Awarenes'' or ``The Global
Foundation for Cultural Recognition'', or else they are shown no name.  These 
names were chosen to signal interest, on the part of the requester, in the
cultural or food-related aspects of the images.

Our primary interest lays in the hysteretic effects of in-task priming.  We
include framing mainly as a point of comparison.  A defensible expectation is
that the framing would produce a much stronger effect, because it clearly 
signals the requester's interest in a specific aspect of image content. 
Remarkably, with respect to the specificity of labels, and their content 
(either food- or culture-orineted), in-task priming produces much stronger 
effects.

The measurement of priming effects, however, traditionally involves a prior
assumption of what those effects will be.  At risk of circularity, but lacking
alternatives, the definition of priming itself traditionally becomes 
operationalized in terms of these effects\cite{}.  We seek to alleviate this 
situation by providing an application-independant definition of priming 
suitable for use in studying the computational properties of HPUs.

In the remainder of the introduction, we review prior work in human 
computation and priming.  In the next section, \textbf{Analytical Framework},
we introduce our definition of priming, before presenting our experimental 
\textbf{Results}, \textbf{Discussion}, and \textbf{Conclusion}.  
We present our \textbf{Methods} in detail at the end of the paper.

\subsection*{Prior Work}
- human computation pioneered by von Ahn and L Dabbish when they demonstrated
	that images could be labeled by people via a game platform
	\cite{von2004labeling}

- Human computation is generally thought of in the context of crowdsourcing,
	which implies soliciting work in the form of an open call to an 
	unstructured pool of workers, general without any longterm engagement and 
	limited ability to screen workers screening of workers. 
	\cite{howe2008crowdsourcing}

- As such, quality assurance has received a lot of attention.
	- varying levels of attention and success \cite{kazai2013analysis}
	- Widely varying reports on accuracy from very negative 
	\cite{marsden2009crowdsourcing}, to positive 
	\cite{alonso2009can, grady2010crowdsourcing}
	- quality control \cite{lease2011quality}
		o filtering 
			- honeypots	\cite{snow2008cheap}
		o averaging
		- \cite{alonso2009can}, \cite{paolacci2010running}

	- training \cite{le2010ensuring}

	- prescreening, even for cultural effects
		\cite{paolacci2010running}
			- worker rep

	- game theoretic treatments
	- worker qualification 					  |	
	- price									  | \cite{kazai2013analysis}
	- task complexity						  |
	- spam, variability, adversarial behavior |

- effects of motivation
	for fun does badly, for money does well
	claimed expertize does badly
	interest does better than boring, and boring performs faster, 
	difficult does worse
	pay 'ok' does best

- studied effects of autonomy, skill use, and meaningfullness, in 
- other studies study the effects of reputation and social standing
- studied effect of giving workflow context\cite{Kinnaird2012281}
- studied effects of task design (user interface): \cite{Finnerty2013}

NOTES ON PRIMING
- when shown an image, a person is better at recognizing the image in a 
	tachistoscope up to 45 min later.\cite{BJOP:BJOP1796}
	* the mechanisms are complex: showing an image of something that had the
		same name also helps recognition, but being presented with a word and
		reading it aloud does not

- priming can be auditory(\cite{BJOP:BJOP1826}) or visual(\cite{BJOP:BJOP1796})

- priming has been detected for words sharing morphemes 
	(reflects vs reflecting) but not for semantically related words that such 
	as held and holding.\cite{BJOP:BJOP1826}
	- suggests stimulus encoding \cite{beller1971priming}\cite{BJOP:BJOP1826}
	\cite{BJOP:BJOP1796} and memory access \cite{beller1971priming} priming 
	effects exist.  Other studies provide definitive evidence that priming
	leads to better application of object knowledge and top-down processes
	\cite{Ghuman17062008}

- existing work suggests that there are separble verbal- and image-based
	semantic systems, as shown through studies of event-related brain 
	potentials in response to high and low imageable words \cite{Swaab200299} 
- priming is an effect whereby previous exposure to a stimulus leads to 
	faster, more accurate responses to similar stimuli \cite{Ghuman17062008}, 
	or lower response thresholds \cite{BJOP:BJOP1826}

@@

\section*{Acknowledgements}
\section*{References}
\begingroup
\renewcommand{\chapter}[2]{}
\bibliography{newbib.bib}
\endgroup
\bibliographystyle{plain}
